fishtraits

### 2019-09-09 data-raw/fishtraits.Rmd
Get id's for fishtraits

Read in files
```{r}
fishtraits <- dir(here::here("inst", "extdata", "fishtraits"), "*.csv") %>%
  paste0(here::here("inst", "extdata", "fishtraits"), "/", .) %>%
  map_dfr(read_csv) %>% 
  filter(id != 810) %>% #one species has a duplicate entry with no data for some reason
  select(scientificName = scientificname, itistsn, commonName = commonname, altcommonName = othernames, 
                      diet_nonfeed = nonfeed, diet_benthic = benthic, diet_surwcol = surwcol, diet_algphyto = algphyto, diet_macvascu = macvascu, 
                      diet_detritus = detritus, diet_inv = invlvfsh, diet_fshcrcrb = fshcrcrb, diet_blood = blood, diet_eggs = eggs, diet_other = other, 
                      max_length = maxtl, matuage, longevity, fecundity, serial, season, starts_with("a_"), starts_with("b_"), starts_with("c1_"),
                      euryhaline, mintemp, maxtemp, muck, claysilt, sand, gravel, cobble, boulder, bedrock, vegetat, debrdetr, lwd,
                      pelagic, preflot, preflen, largeriv, smallriv, creek, sprgsubt, lacustrine, potanadr, lowland, upland, montane,
                      slowcurr, modcurr, fastcurr)

#find species without a TRUE habitat type 
missing_hab <- fishtraits %>% 
  #select(starts_with("a_"), starts_with("b_")) %>%
  filter_at(vars(starts_with("a_"), starts_with("b_"), starts_with("c1_")),  ~ !. %in% TRUE) %>%
  pull(itistsn) %>%
  unique()

fishtraits_clean <- fishtraits %>%
  gather(habitat, value, starts_with("a_"), starts_with("b_"), starts_with("c1_")) %>%
  mutate(habitat = case_when(
    itistsn %in% missing_hab ~ NA_character_,
    TRUE ~ habitat
    )
  ) %>% 
  filter(value == TRUE | is.na(habitat), itistsn != 0) %>%
  select(-value) %>%
  distinct()

@@

>>>>

Read in files
```{r}
fishtraits_data <- dir(here::here("inst", "extdata", "fishtraits"), "*.csv") %>%
  paste0(here::here("inst", "extdata", "fishtraits"), "/", .) %>%
  map_dfr(read_csv) %>% 
  filter(id != 810) %>% #one species has a duplicate entry with no data for some reason
  select(scientificName = scientificname, itistsn,
                      diet_nonfeed = nonfeed, diet_benthic = benthic, diet_surwcol = surwcol, diet_algphyto = algphyto, diet_macvascu = macvascu, 
                      diet_detritus = detritus, diet_inv = invlvfsh, diet_fshcrcrb = fshcrcrb, diet_blood = blood, diet_eggs = eggs, diet_other = other, 
                      max_length = maxtl, matuage, longevity, fecundity, serial, season, starts_with("a_"), starts_with("b_"), starts_with("c1_"),
                      euryhaline, mintemp, maxtemp, muck, claysilt, sand, gravel, cobble, boulder, bedrock, vegetat, debrdetr, lwd,
                      pelagic, preflot, preflen, largeriv, smallriv, creek, sprgsubt, lacustrine, potanadr, lowland, upland, montane,
                      slowcurr, modcurr, fastcurr)

#find species without a TRUE habitat type 
missing_hab <- fishtraits_data %>% 
  #select(starts_with("a_"), starts_with("b_")) %>%
  filter_at(vars(starts_with("a_"), starts_with("b_"), starts_with("c1_")),  ~ !. %in% TRUE) %>%
  pull(itistsn) %>%
  unique()

fishtraits_clean <- fishtraits_data %>%
  gather(habitat, value, starts_with("a_"), starts_with("b_"), starts_with("c1_")) %>%
  mutate(habitat = case_when(
    itistsn %in% missing_hab ~ NA_character_,
    TRUE ~ habitat
    )
  ) %>% 
  filter(value == TRUE | is.na(habitat), itistsn != 0) %>%
  select(-value) %>%
  distinct() %>%
  mutate_at(vars(-c(scientificName, itistsn, max_length, matuage, longevity, fecundity, 
                    season, mintemp, maxtemp, habitat)), ~ as.integer(.x))

@@

Get id's! They seemingly already have ITIS id's, but let's double check that they're up to date with what we have through taxadb
```{r}
fishtraits <- fishtraits_clean %>%
  mutate(itistsn = paste0("ITIS:", itistsn), 
         id = get_ids(scientificName))

#There are no instances where the database id doesn't match the id from taxadb
fishtraits %>% filter(itistsn != id) %>% dim()

#so we'll just keep that ID column
fishtraits <- fishtraits %>%
  mutate(id = itistsn) %>%
  select(-itistsn)
```

```{r}
usethis::use_data(fishtraits)
```
